---
title: "Restaurant"
author: "Oriel Zambrano"
date: "December 9, 2016"
output: html_document
---


This is an  analysis for  [the code challenge](http://https://github.com/TOTVS/MDMStatic/tree/master/code-challenge) forecast

Load and clean data

```{r, echo=TRUE, warning=FALSE}
setwd("~/Documents/Oriel/Notas fiscais")
#LOAD DATA
#first  i changed txt file to json file only modifing the extension
#read text
library(jsonlite)
library(dplyr)
library(ggplot2)
library(lubridate)
library(tidyverse)
library(scales)
library(ggthemes)
jsonfile<-readLines("sample.json",skipNul = T)
jsontovts<-fromJSON(paste(jsonfile, collapse = " "))

#Get sales

receita<-cbind(as.numeric(rownames(jsontovts$complemento)), jsontovts$complemento)
names(receita)<- c("id", "vrTotal")

#get and imput purchase dates  

dates<-as.data.frame(jsontovts$ide$dhEmi$`$date`)
names(dates)<-"purchasedatehour"
receita<-cbind(receita,dates)
#convert to date type and add time

receita$purchasedate<-as.Date(receita$purchasedatehour, format="%Y-%m-%dT%H:%M")
receita$purchasedatehour<-as.POSIXct(receita$purchasedatehour, format="%Y-%m-%dT%H:%M", tz = "GMT")
receita$time<-format(as.POSIXct(receita$purchasedatehour, format="%Y-%m-%d %H:%M"), format="%H:%M")
receita$hour<-as.numeric(format(as.POSIXct(receita$purchasedatehour, format="%Y-%m-%d %H:%M"), format="%H"))
receita$minutes<-as.numeric(format(as.POSIXct(receita$purchasedatehour, format="%Y-%m-%d %H:%M"), format="%M"))
receita$weekday<- weekdays(receita$purchasedate, abbreviate = T)
receita$wday<-as.numeric(format(as.Date(receita$purchasedate),"%w"))
receita$weeknum<-isoweek(ymd(receita$purchasedate))  


#what table number
receita$mesa<-as.numeric(gsub(jsontovts$infAdic$infCpl, pattern = "Mesa", replacement =""))
  
#mealtype  dinner or lunch Saturdays

for(i in 1:length(receita$hour)) { 
  if(receita$hour[i]<17) {receita$meal[i]<-"lunch"} else {receita$meal[i]<-"dinner"}
}

#extract data from orders.

dflist<-jsontovts$dets
for(i in 1:length(dflist)) { 
  dflist[[i]]$prod$id<-i}  #introduce ID  for each order
dfto<-list()
for(i in 1:length(dflist)) {
  dfto[[i]]<-dflist[[i]]$prod
}

ordersdf<-do.call(rbind, dfto) #create a data.frame

sort(table(ordersdf$xProd), decreasing = T) #everybody ate buffet. So buffet price is sensible and important.


productorder<-ordersdf %>% group_by(id,xProd) %>% summarise(qCom=sum(qCom)) %>% spread(xProd, qCom)
productvalue<-ordersdf %>% group_by(id,xProd) %>% summarise(vProd=sum(vProd)) %>% spread(xProd, vProd)

names(productorder)<- c("id"  ,           "AGUA"     ,      "BACARDI"  ,      "BUFFET"     ,    "BULE_CHA"    ,  "CAFE_EXPRESSO" ,
"CAIPIRINHA"  ,   "CAIPIROSKA"   ,  "CERVEJA"     ,   "CERVEJA_LATA"  , "CHA"      ,      "DOCINHOS"      ,
"HARUMAKI"    ,   "LIMONADA"     ,  "REFRIGERANTE"  , "SAKE"        ,   "SASHIMI"  ,      "SOBREMESA"    , 
 "SUCO"        ,   "SUSHI_ESPECIAL" ,"TEMAKI"     ,    "URAMAKI"    ,    "VINHO"      ,    "WHISKY"      ,  
"YAKISSOBA"   )  
names(productvalue)<- c("id"  ,           "AGUA"     ,      "BACARDI"  ,      "BUFFET"     ,    "BULE_CHA"    ,  "CAFE_EXPRESSO" ,
                        "CAIPIRINHA"  ,   "CAIPIROSKA"   ,  "CERVEJA"     ,   "CERVEJA_LATA"  , "CHA"      ,      "DOCINHOS"      ,
                        "HARUMAKI"    ,   "LIMONADA"     ,  "REFRIGERANTE"  , "SAKE"        ,   "SASHIMI"  ,      "SOBREMESA"    , 
                        "SUCO"        ,   "SUSHI_ESPECIAL" ,"TEMAKI"     ,    "URAMAKI"    ,    "VINHO"      ,    "WHISKY"      ,  
                        "YAKISSOBA"   )  


productorder[is.na(productorder)]<-0
productvalue[is.na(productvalue)]<-0

qcomandprodctvalue<-merge(productorder,productvalue, by = "id",suffixes = c(".qCom", ".Vpaid"))

#merge in the big dataframe
receita<-merge(x =receita, y =qcomandprodctvalue, all.x = T, all.y = F)
```




There are no important difference between average sales by table code. But tables 1 to 50 sells more than tabeles 51 to 99..( tehere are some wierd tables never used, like 12, 20, 50)


```{r, echo=TRUE}
receita %>%group_by(mesa) %>% summarise(sales=sum(vrTotal)) %>% mutate(CumulativeSales=cumsum(sales)) %>% 
  ggplot()+geom_point(aes(x = mesa, y = CumulativeSales)) + labs(title= "Cumulative sales by table code", x= "Table Code", y ="Sales") +
  theme_few() +scale_x_continuous(breaks = seq(0,100,by = 5))
```

##Pattern

he restaurant is opening each week a new dinner service. first week, just monaday and tuesday, second week, 3 days  for dinner, third week 4 days for dinner. ( We can think tha the next week they will open friday to) 
(because year is beginnig)


```{r}
receita %>% group_by(purchasedate, meal) %>% summarise(sales=sum(vrTotal)) %>% arrange(desc(meal)) %>% mutate(pos=cumsum(sales)-0.5*sales) %>%
  ggplot() + geom_bar(aes(x = purchasedate, y = sales, fill=meal, label=sales), stat = "identity") +
  geom_text(aes(x = purchasedate, y = pos, fill=meal, label=dollar(sales)), size=3) + scale_y_continuous(labels=dollar_format()) +labs(title="Sales by day and shift", x="Date", y="Totalsales")
```



##Correlations

There are no strong relationship on features, just buffet cunsumption and sales the other featues have no strong relationship and ore have obvious relations. Like when consume whisky there are more sales for the obvious reason that whisky is spensive.

```{r, echo=TRUE, fig.height=15, fig.width=15}
#CORPPLOT
corrmatrix<-receita %>% select(-c(weekday,time, meal,purchasedatehour,purchasedate) ) %>% cor()
library(corrplot)
corrplot::corrplot(corrmatrix)
```


##Forecast


The next week (week 4) there will be sales for  $ 39 281

```{r, echo=TRUE}
#Naive Lunch forecast
dflunch<-receita %>% group_by(weeknum=as.numeric(weeknum),weekday, purchasedate, meal) %>% arrange(desc(meal)) %>% summarise(total=sum(vrTotal))
library(caret)

#randomforest
control <- trainControl(method="repeatedcv", number=6, repeats=4, search="grid")
set.seed(1254)
tunegrid <- expand.grid(.mtry=3, .ntree=500)
library(randomForest)

modellinardflunchrf <-train(total~weeknum+weekday+meal,data = dflunch, method="rf", metric="RMSE", ntree=1000, trControl=control)

#create empty dates and shifts
dflunch[27:37,1]<- rep(4)
dflunch[27:37,2]<- c("Seg","Ter","Qua","Qui","Sex","Sáb","Seg","Ter","Qua","Qui","Sex")
dflunch[27:32,3] <- seq(as.Date("2016-01-25"), as.Date("2016-01-30"), by = "day")
dflunch[33:37,3] <- seq(as.Date("2016-01-25"), as.Date("2016-01-29"), by = "day")
dflunch[27:32,4]<-rep("lunch")
dflunch[33:37,4]<-rep("dinner")  


#forecast random forest
dflunch[27:37,5]<- predict(modellinardflunchrf, dflunch[27:37,])

#Plot Everything
dflunch %>% group_by(purchasedate, meal) %>% summarise(sales=sum(total)) %>% arrange(desc(meal)) %>% mutate(pos=cumsum(sales)-0.5*sales) %>%
  ggplot() + geom_bar(aes(x = purchasedate, y = sales, fill=meal, label=sales), stat = "identity") +
  geom_text(aes(x = purchasedate, y = pos, fill=meal, label=dollar(sales)), size=3) + scale_y_continuous(labels=dollar_format()) +labs(title="Sales by day and shift", x="Date", y="totalsales")

```

##forecast orders

```{r}
#Forecast people at the restaurant

dforders<-receita %>% group_by(weeknum=as.numeric(weeknum),weekday, purchasedate, meal) %>% arrange(desc(meal)) %>% summarise(totalorders=n_distinct(id))

#randomforest
control <- trainControl(method="repeatedcv", number=6, repeats=4, search="grid")
tunegridFOOD<- expand.grid(.mtry=c(3,4,5))
modellinardflunchrf <-train(totalorders~weeknum+weekday+meal,data = dforders, method="rf", metric="RMSE", ntree=500, trControl=control, tuneGrid=tunegridFOOD)

#create empty dates and shifts
dforders[27:37,1]<- rep(4)
dforders[27:37,2]<- c("Seg","Ter","Qua","Qui","Sex","Sáb","Seg","Ter","Qua","Qui","Sex")
dforders[27:32,3] <- seq(as.Date("2016-01-25"), as.Date("2016-01-30"), by = "day")
dforders[33:37,3] <- seq(as.Date("2016-01-25"), as.Date("2016-01-29"), by = "day")
dforders[27:32,4]<-rep("lunch")
dforders[33:37,4]<-rep("dinner")  


#random forest
dforders[27:37,5]<- predict(modellinardflunchrf, dforders[27:37,])

dforders %>% group_by(purchasedate, meal) %>% summarise(orders=sum(totalorders)) %>% arrange(desc(meal)) %>% mutate(pos=cumsum(orders)-0.5*orders) %>%
  ggplot() + geom_bar(aes(x = purchasedate, y = orders, fill=meal, label=orders), stat = "identity") +
  geom_text(aes(x = purchasedate, y = pos, fill=meal, label=round(orders)), size=3)
```


```{r, echo=TRUE}
#Forecast orders
dforders[27:37,] %>% arrange(purchasedate) %>% print(n=30)
dforders %>% group_by(weeknum) %>% summarise(orders=round(sum(totalorders)))

```


##Forecast food, sodas, and aguas.

Just for buffet, sodas and water, are the products with more stock attention. 

```{r, echo=TRUE}
#QUANTITY OF FOOD

dffood<-receita %>% group_by(weeknum=as.numeric(weeknum),weekday, purchasedate, meal) %>% arrange(desc(meal)) %>% summarise(totalfood=sum(BUFFET.qCom))
#randomforest
control <- trainControl(method="repeatedcv", number=6, repeats=4, search="grid")
set.seed(1254)
modellinardflunchrf <-train(totalfood~weeknum+weekday+meal,data = dffood, method="rf", metric="RMSE", ntree=300, trControl=control)
#create empty dates and shifts
dffood[27:37,1]<- rep(4)
dffood[27:37,2]<- c("Seg","Ter","Qua","Qui","Sex","Sáb","Seg","Ter","Qua","Qui","Sex")
dffood[27:32,3] <- seq(as.Date("2016-01-25"), as.Date("2016-01-30"), by = "day")
dffood[33:37,3] <- seq(as.Date("2016-01-25"), as.Date("2016-01-29"), by = "day")
dffood[27:32,4]<-rep("lunch")
dffood[33:37,4]<-rep("dinner")  

#forecast random forest
dffood[27:37,5]<- predict(modellinardflunchrf, dflunch[27:37,])

dffood %>% group_by(purchasedate, meal) %>% summarise(totalfood=sum(totalfood)) %>% arrange(desc(meal)) %>% mutate(pos=cumsum(totalfood)-0.5*totalfood) %>%
  ggplot() + geom_bar(aes(x = purchasedate, y = totalfood, fill=meal, label=totalfood), stat = "identity") +
  geom_text(aes(x = purchasedate, y = pos, fill=meal, label=round(totalfood)), size=3) +labs(title="Kilos of food consumed and forecasted", x="Date", y="Kilograms of food")
```

The next week there will be consumed 473 kilos of food.

```{r, echo=TRUE}
#Forecast Buffet
dffood[27:37,] %>% arrange(purchasedate) %>% print(n=30)

dffood %>% group_by(weeknum) %>% summarise(food=(sum(totalfood)))
```

```{r}
#Forecast for Refrigerante (soda)

dfrefri<-receita %>% group_by(weeknum=as.numeric(weeknum),weekday, purchasedate, meal) %>% arrange(desc(meal)) %>% summarise(totalrefri=sum(REFRIGERANTE.qCom))
#randomforest
control <- trainControl(method="repeatedcv", number=6, repeats=4, search="grid")
set.seed(1254)
modellinardflunchrf <-train(totalrefri~weeknum+weekday+meal,data = dfrefri, method="rf", metric="RMSE", ntree=300, trControl=control)
#create empty dates and shifts
dfrefri[27:37,1]<- rep(4)
dfrefri[27:37,2]<- c("Seg","Ter","Qua","Qui","Sex","Sáb","Seg","Ter","Qua","Qui","Sex")
dfrefri[27:32,3] <- seq(as.Date("2016-01-25"), as.Date("2016-01-30"), by = "day")
dfrefri[33:37,3] <- seq(as.Date("2016-01-25"), as.Date("2016-01-29"), by = "day")
dfrefri[27:32,4]<-rep("lunch")
dfrefri[33:37,4]<-rep("dinner")  

#forecast random forest
dfrefri[27:37,5]<- predict(modellinardflunchrf, dfrefri[27:37,])

dfrefri %>% group_by(purchasedate, meal) %>% summarise(totalrefri=sum(totalrefri)) %>% arrange(desc(meal)) %>% mutate(pos=cumsum(totalrefri)-0.5*totalrefri) %>%
  ggplot() + geom_bar(aes(x = purchasedate, y = totalrefri, fill=meal, label=totalrefri), stat = "identity") +
  geom_text(aes(x = purchasedate, y = pos, fill=meal, label=round(totalrefri)), size=3) + scale_y_continuous() +labs(title="Number of sodas consumed and forecasted", x="Date", y="Number of sodas")

```
The next week there will be consumed 569 sodas.

```{r}
#Forecast soda
dfrefri[27:37,] %>% arrange(purchasedate) %>% print(n=30)

dfrefri %>% group_by(weeknum) %>% summarise(refri=(sum(totalrefri)))
```


```{r}
dfagua<-receita %>% group_by(weeknum=as.numeric(weeknum),weekday, purchasedate, meal) %>% arrange(desc(meal)) %>% summarise(totalagua=sum(AGUA.qCom))
#randomforest
control <- trainControl(method="repeatedcv", number=6, repeats=4, search="grid")
set.seed(1254)
modellinardflunchrf <-train(totalagua~weeknum+weekday+meal,data = dfagua, method="rf", metric="RMSE", ntree=100, trControl=control)
#create empty dates and shifts
dfagua[27:37,1]<- rep(4)
dfagua[27:37,2]<- c("Seg","Ter","Qua","Qui","Sex","Sáb","Seg","Ter","Qua","Qui","Sex")
dfagua[27:32,3] <- seq(as.Date("2016-01-25"), as.Date("2016-01-30"), by = "day")
dfagua[33:37,3] <- seq(as.Date("2016-01-25"), as.Date("2016-01-29"), by = "day")
dfagua[27:32,4]<-rep("lunch")
dfagua[33:37,4]<-rep("dinner")  

#forecast random forest
dfagua[27:37,5]<- predict(modellinardflunchrf, dfagua[27:37,])

dfagua %>% group_by(purchasedate, meal) %>% summarise(totalagua=sum(totalagua)) %>% arrange(desc(meal)) %>% mutate(pos=cumsum(totalagua)-0.5*totalagua) %>%
  ggplot() + geom_bar(aes(x = purchasedate, y = totalagua, fill=meal, label=totalagua), stat = "identity") +
  geom_text(aes(x = purchasedate, y = pos, fill=meal, label=round(totalagua)), size=3) + scale_y_continuous() +labs(title="Number of water bottles consumed and forecasted", x="Date", y="Number of water bottles")

```

The next week there will be consumed 293 sodas.

```{r}
dfagua[27:37,] %>% arrange(purchasedate) %>% print(n=30)

dfagua %>% group_by(weeknum) %>% summarise(aguas=(sum(totalagua)))
```


```{r}
finaldataforecasted<-data.frame(dflunch$purchasedate, 
                   dflunch$meal,  dflunch$total, dforders$totalorders, dffood$totalfood, dfrefri$totalrefri, dfagua$totalagua)

print(finaldataforecasted)

```

